---
title: "class 12"
author: "Michelle Louie"
date: "5/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("bio3d")
file.name <- get.pdb("1hsg")
hiv <- read.pdb(file.name)
hiv
```

Q1: What is the name of the two non protein resid values in this structure?  HOH (127), MK1 (1)]

What does resid correspond to and how would you get a listing of all reside values in this structure?



```{r}
# select out protein only and ligind only
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")

# write out these new pdb files so we can use them for docking 
write.pdb(prot, file="1hsg_protein.pdb")
write.pdb(lig, "1hsg_ligand.pdb")
```

## processing docking results for vmd display 

```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")

ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```
 
 the 1st result is the best fit frame at which the ligand position almost exactly alignes with the MK1 position 
 
 # rmsd of our results 'res' tp <arck compound

select non-hydrogen atoms from our docking results 
 
```{r}
inds.res <- atom.select(res, "noh")
inds.res

res$xyz[ ,inds.res$xyz]
```
 
```{r}
rmsd(lig, res$xyz[ ,inds.res$xyz])
```
 
