---
title: "Class 11"
author: "Michelle Louie"
date: "5/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB STATS

Download CSV file from PDB database <>, read this into R and determine fraction of X -reay structure. 
```{r}
pdb.stats <- read.csv("Data Export Summary.csv")
```

Lets calculate the fraction

```{r}
percent <- ( pdb.stats$Total / sum(pdb.stats$Total) ) *100
names(percent) <- pdb.stats$Experimental.Method 
percent 
```

Q1: What proportion of PDB entries does X-ray crystallography account for? What proportion of structures are protein?

X-Ray is 89.5% 

Q2: Type HIV in the search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

6

## Introduction to Bio3D in R

loading the Bio3D package

```{r}
# intsall.packages("bio3d)"
library(bio3d)
```


list function of package and try an example
```{r}
lbio3d()
example(plot.bio3d)
```


```{r}
pdb<- read.pdb("1hsg")
pdb
```

select and write a c-alpha atoms only pdb 
```{r}
inds <- atom.select( pdb, "calpha")
inds
# 198 c-alpha atoms selected 
# c-alpha = the backbone of protein :3
```

checking if e have what we want
```{r}
head( pdb$atom[inds$atom, ] )
```

looks good. Let's continue to trim pdb and write it out 
```{r}
ca.pdb <- trim.pdb( pdb, inds=inds)
ca.pdb
write.pdb (ca.pdb, file="calpha_only.pdb")
```

## working with multiple PDB files 

```{r}
# Download some example PDB files
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)
```

```{r}
 # Extract and align the chains we are interested in
pdbs <- pdbaln(files, fit = TRUE)

# Print to screen a summary of the 'pdbs' object
pdbs
```

```{r}
# Access the first 5 rows, and 8 columns
pdbs$ali[1:5, 1:8]

# Associated residues numbers
pdbs$resno[1:5, 1:8]
```

```{r}
# Inspect alignment for gaps
gaps <- gap.inspect(pdbs$ali)

# Indices of non-gap columns
head(gaps$f.inds)

# Access non-gap columns
pdbs$ali[, gaps$f.inds]
```

```{r}
# Calculate sequence identity
seqidentity(pdbs)

# Calculate RMSD (root mean square deviation)
rmsd(pdbs)
```

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Clustering
hc <- hclust(as.dist(rd))
grps <- cutree(hc, k=3)

# Plot results as dendrogram
hclustplot(hc, k=3)
```

```{r}
# Perform PCA
pc <- pca(pdbs)
plot(pc, col=grps)
```

## Example Application on Adenylate Kinase (Adk)

find sequence
```{r}
aa <- get.seq("1ake_A")
aa
## Fetching... Please wait. Done.
```

search the pdb database
```{r, cache=TRUE}

# Blast or hmmer search
b <- blast.pdb(aa)
## Searching ... please wait (updates every 5 seconds) RID = F3XPYPMH01R 
## .
## Reporting 90 hits
```
plot
```{r}
# plotting the -log (e-values)
hits <- plot(b)
hits
```

```{r}
head(hits$pdb.id)
## [1] "1AKE_A" "1AKE_B" "1ANK_A" "1ANK_B" "4AKE_A" "4AKE_B"
```

```{r}
# Align structures
pdbs <- pdbaln(files)

# Vector containing PDB codes
ids <- basename.pdb(pdbs$id)

# Draw schematic alignment 
plot(pdbs, labels=ids)
```

```{r}
# Calculate sequence conservation
cons <- conserv(pdbs, method="entropy22")

# SSE annotations
sse <- pdbs2sse(pdbs, ind=1, rm.gaps=FALSE)

# Plot conservation per residue
plotb3(cons, sse=sse, ylab="Sequence entropy")
  
```

