---
title: "class 16"
author: "Michelle Louie"
date: "5/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TnSeq_practice

```{r}
# install.packages("tidyverse")

library(tidyr)
library(dplyr)
library(ggplot2)
library(readr)
```
```{r}

exp_data<-read.csv("Experimental_results.csv")
head(exp_data)
```

```{r}
# remove column X from the dataframe:
exp_data <- select(exp_data, -X)
head(exp_data)
```

```{r}
# select() pratice 

Mut_str_env <- select(exp_data, Mut_ID, Strain, Environment)
head(Mut_str_env)
```

Information from a ggplot description page:
**ggplot(df, aes(x, y, <other aesthetics>))**

We are going to plot freq of barcode over time, but our df isnt set up properly to be able to define x,y.

Lets rearrange our table to be able to plot the data easily. Instead on keeping information about barcode frequency in rows, we are going to create a column “Time” with time points and a column “Frequency” with corresponding barcode frequencies.

```{r}
# First, check how function "gather" works
exp_rearranged <- gather(exp_data, Generation, Frequency,H_0:H_96)
head(exp_rearranged)
```

You might have noticed that “Generation” column contains both “H” that stands for “hours” and numbers. Lets remove “H_” part from this column.

Check the syntax of “separate” function.
```{r}
# Separate values in "Generation" column into 2 columns

table_for_graph <-separate(exp_rearranged,Generation,into=c("H","Time"))

head(table_for_graph)
```

```{r}
# Remove column "H" using function "select"
table_for_graph <- select(table_for_graph, -H)
table_for_graph
```
```{r}
table_cleaned <- na.omit(table_for_graph)
table_cleaned$Time <- as.numeric(table_cleaned$Time)
head(table_cleaned)
```

We can graph them with different color to compare them, but the following way has too much overlap and hard to read ;<
```{r}
# We might need to plot data for each strain separately..
DivAnc <- filter(table_cleaned, table_cleaned$Strain=="DivAnc")
L013 <- filter(table_cleaned, table_cleaned$Strain=="L013")
```

```{r}
# make a plot for DivAnc strain
DivAnc_plot=ggplot(DivAnc)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#000033")+ggtitle("DivAnc_SC3")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
DivAnc_plot
```

```{r}
# make a plot for L013 strain
L013_plot=ggplot(L013)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#CC6633")+ggtitle("L013_SC3")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
L013_plot
```

Can we make 2 graphs at the same time?
```{r}
ggplot(table_cleaned)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#000033")+facet_grid(.~Strain)+ggtitle("Barcode trajectories")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
```

Lets pick one mutation and check how it behaves in different strains

```{r}
mut34<-filter(table_cleaned, table_cleaned$Mut_ID=="34")
mut34       
```

```{r}
ggplot(mut34,aes(Time, Frequency, group=BC_ID, color=BC_ID))+geom_line()+theme(legend.position="none")+facet_grid(.~Strain)+ggtitle("Mutation_34")+xlab("Time, hours") + ylab("Log10(Barcode frequency)")+theme(plot.title = element_text(hjust = 0.5))
```
let's fliter out the data 
```{r}
mut34_f<-filter(mut34, mut34$Frequency>(-5))
mut34_f
```

Plot again the same type of graph, but use filtered data.
```{r}
ggplot(mut34_f,aes(Time, Frequency, group=BC_ID, color=BC_ID))+geom_line()+theme(legend.position="none")+facet_grid(.~Strain)+ggtitle("Mutation_34")+xlab("Time, hours") + ylab("Log10(Barcode frequency)")+theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(mut34_f,aes(Time, Frequency, colour = BC_ID, group=BC_ID)) + geom_point() + geom_smooth(se = FALSE, method = "lm")+facet_grid(.~Strain)+theme(legend.position="none")+ggtitle(paste("Mutation",34, sep="_"))+xlab("Time, hours")+ ylab("Log10(Barcode frequency)")
```

Now it’s time to estimate slope for each barcode. Lets greate a file that will contain information about BC_ID, Mut_ID, Strain, and estimated slope.

```{r}
# Lets become familiar with lm function:

# For this exercise, take the filtered data for mutation 34 (mut34_f) and filter out information about one barcode you like

# I have chosen BC_ID=25361 in DivAnc strain
BC_25361<-filter(mut34_f, mut34_f$BC_ID=="25361", mut34_f$Strain=="DivAnc")
BC_25361
```

```{r}
#Lets plot frequency of this barcode:
barcode_plot=ggplot(BC_25361,aes(Time, Frequency, colour = BC_ID))+geom_point()+theme(legend.position="none")+ggtitle("BC_25361")+xlab("Time, hours") + ylab("Log10(Frequency)")
```
```{r}
#Lets use lm function to fit the line to these points:
ggplot(BC_25361,aes(Time, Frequency, colour = BC_ID))+geom_point()+geom_smooth(se = FALSE, method = "lm")+theme(legend.position="none")+ggtitle("BC_25361")+xlab("Time, hours") + ylab("Log10(Frequency)")

#alt way of writing the same thing 
barcode_plot_ln=barcode_plot+geom_smooth(se = FALSE, method = "lm")
barcode_plot_ln
```
```{r}
# Lets check what data does lm function return:
regression_model<-lm(Frequency~Time,BC_25361)
summary_data<-summary(regression_model)
summary_data
```

```{r}
# The information we are interested in is the value of Slopeand Intercept of this line:
# Let's try to access them:

# Time
Time<-summary_data$coefficients[2]
Time
```

```{r}
# Lets create the file:
data_header=matrix(data = NA,nrow = 1,ncol = 7)
        data_header[1]="Mut_ID"
        data_header[2]="BC_ID"
        data_header[3]="Strain"
        data_header[4]="Slope"
        data_header[5]="Intercept"
        data_header[6]="R^2"
write.table(data_header,"~/Desktop/BIMM143/Tnseq_practice_output.csv",append = FALSE, sep = ",",eol="\n",dec=".",row.names = FALSE,col.names = FALSE)
```

```{r}
for (mut in unique(table_cleaned$Mut_ID)) {
    mut_data=filter(table_cleaned,table_cleaned$Mut_ID==paste(mut))
    #now we have all data for each mutation separately
    for (bc in unique (mut_data$BC_ID)) {
      #now we filtered data for each barcode within 1 mutation
      bc_mut_data=filter(mut_data,mut_data$BC_ID==paste(bc))
      for (strain in unique (bc_mut_data$Strain)) {
        str_bc_mut_data=filter(bc_mut_data,bc_mut_data$Strain==paste(strain))
        #only considering combinations with 3 or more data points - anything less is statistically insignificant
        if (nrow(str_bc_mut_data)>2){
          regression_model=lm(Frequency~Time,str_bc_mut_data)
          summary_data=summary(regression_model)
          #now write to the output file! Prepare the data array first
          data_output=matrix(data = NA,nrow = 1,ncol = 6)
          data_output[1]=mut
          data_output[2]=bc
          data_output[3]=strain
          #slope
          data_output[4]=summary_data$coefficients[2]
          #intercept
          data_output[5]=summary_data$coefficients[1]
          #r-squared
          data_output[6]=summary_data$r.squared
          #time to write
          write.table(data_output,"~/Desktop/BIMM143/Tnseq_practice_output.csv",append = TRUE, sep = ",",eol="\n",dec=".",row.names = FALSE,col.names = FALSE)
      }
    }
  }
 }
```

