---
title: "CLASS 13"
author: "Michelle Louie"
date: "5/15/2018"
output: 
  html_document:
    keep_md: yes 

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Q1: What are those 4 candidate SNPs?
rs12936231, rs8067378, rs9303277, and rs7216389

Q2: What three genes do these variants overlap or effect?
ZPBP2, GSDMB, and ORMDL3

Q3: What is the location of rs8067378 and what are the different alleles for rs8067378?
Chromosome 17: 39895095 (forward strand) A/G variants (43% G)

Q4: What are the downstream genes for rs8067378? Any genes named ZPBP2, GSDMB, and ORMDL3


## 1000 genomes data analysis 
 
download genotype data from Ensemble for childhood asthma associated SNP rs8067378 
see: (MXL) Mexican Ancestry in Los Angeles, California

```{r}
# read.csv
genotype <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(genotype)
```

find how many of each genotype there are in this dataset 
```{r}
genotype.summary <- table(genotype[ ,2])
genotype.summary
```

proportion of each genotype
```{r}
genotype.summary/ nrow(genotype) * 100
```

> **Q5:** What proportion of the Mexican Ancestry in Los Angeles sample population (MXL) are homozygous for the asthma associated SNP (G|G)?
Answer: 14%

## RNA seq 
look at sample HG00109

> **Q6:** Back on the ENSEMBLE page, search for the particular sample HG00109. This is a male from the GBR population group. What is the genotype for this sample?
G|G

```{r}
# install.packages("seqinr")
# install.packages("gtools")

library(seqinr)
library(gtools)

phred<- asc( s2c("DDDDCDEDCDDDDBBDDDCC") )-33
phred
```

```{r}
prob <- 10**(-phred/10)
prob
```

## RNA-seqpop scale analysis

read our result files feom RNA seq across ~230 samples/ppl

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
```

look at data 
```{r}
table(expr$geno)
```

```{r}
#how many sapmles?
nrow(expr)

# Box plot 
boxplot(exp ~ geno, data=expr, notch=TRUE)

install.packages("ggplot2")
library(ggplot2)
ggplot(expr, aes(exp, fill= geno)) + geom_density(alpha=0.2)
```









