---
title: "class 14"
author: "Michelle Louie"
date: "5/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bioconductor setup for R
```{r}
 source("http://bioconductor.org/biocLite.R")
 biocLite()
 # For this class, you'll also need DESeq2:
 bioclite("DESeq2")
```

# Import countData and colData into R
```{r}

counts <- read.csv("data_airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("data_airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
```

```{r}
head(metadata)
```

# Toy differential gene expression
```{r}
View(metadata)
```
```{r}
# looking at metadata's dex column and selecting only the rows labeled "control"
control <- metadata[metadata[ ,"dex"]=="control", ]
control

# access the col correlated with control in count data 
control.mean <- rowSums( counts[ ,control$id] )/ length(control$id) 
names(control.mean) <- counts$ensgene
control.mean 
```

```{r}
head(control.mean) 
```

```{r}
treated <- metadata[metadata[ ,"dex"]=="treated", ]
treated

treated.mean <- rowSums( counts[ ,treated$id] )/ length(treated$id) 
names(treated.mean) <- counts$ensgene
head(treated.mean)
```
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

Plot our data with control.mean as x-axis & treated.mean as y-axis
```{r}
plot(control.mean, treated.mean)
```
edit the scale of the axises to see more data pts b/c they're overlapping in the bottom left corner 
```{r}
plot(control.mean, treated.mean, log="xy")
```

here, we calculate log2foldchange, and add it to our meancount data.frame
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```
 NaN (“not a number””) and -Inf (negative infinity) results.
 
```{r}
# framework example on how to remove the 0 
x <- matrix( c(1:10), ncol=2, byrow=1)
x[5,2] <-0
x
which( x==0, arr.ind=TRUE)
```

```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)
to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```
 
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```
 
```{r}
table(up.ind)
# we want to count the # of TRUE 
paste("Number of UP genes:", sum(up.ind))
paste("Number of DOWN genes:", sum(down.ind))

```


```{r}
anno <- read.csv("annotables_grch38.csv")
head(anno)
```

# Adding annotation data
```{r}
results<- merge(mycounts,anno, by.x= "row.names", by.y="ensgene")
results
```

```{r}
biocLite("AnnotationDbi")
library("AnnotationDbi")
biocLite("org.Hs.eg.db"n)
library("org.Hs.eg.db")
```
```{r}
columns(org.Hs.eg.db)
```

```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
head(mycounts)
```

# DESeq2 analysis

load the package
```{r}
library(DESeq2)
citation("DESeq2")
```

construct a DESeqDataSet object from (1) a count matrix, (2) a metadata file, and (3) a formula indicating the design of the experiment
```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

Run the calculations with DESeq()
```{r}
dds <- DESeq(dds)
```
view the results 
```{r}
res<- results(dds)
res
```

```{r}
summary(res)
```
```{r}
# order our results table by the smallest p value
resOrdered <- res[order(res$pvalue),]
```

customize the results table
```{r}
#  alpha should be set to the adjusted p value cutoff
res05 <- results(dds, alpha=0.05)
summary(res05)
```

accessing the actual subset of the data.frame passing a threshold 
```{r}
# How many are significant with an adjusted p-value < 0.05
resSig05 <- subset(as.data.frame(res), padj < 0.05)
nrow(resSig05)

#How many are significant with an adjusted p-value < 0.01
resSig01 <- subset(as.data.frame(res), padj < 0.01)
nrow(resSig01)
```

arrange and view the results by the adjusted p-value
```{r}
ord <- order( resSig01$padj )
#View(res01[ord,])
head(resSig01[ord,])
```

**Q10.** Using either the previously generated anno object (annotations from the file annotables_grch38.csv file) or the mapIds() function (from the AnnotationDbi package) add annotation to your res01 results data.frame.
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
columns(org.Hs.eg.db)
```

```{r}
resSig01$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(resSig01),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
head(resSig01)
```

```{r}
write.csv(resSig01[ord,], "signif01_results.csv")
```

sort() vs order()
```{r}
x<- c(1, 21, 3, 6)
# will sort the element in vector by low to high #
sort(x)

# will give the position if you were to order the element from low to high 
order(x)
```

# Data Visualization

Let’s first see what the gene ID is for the CRISPLD2 gene using:
```{r}
i <- grep("CRISPLD2", resSig01$symbol)
resSig01[i,]

rownames(resSig01[i,])
```

with that gene ID, let’s plot the counts, where our intgroup, or “interesting group” variable is the “dex” column
```{r}
plotCounts(dds, gene="ENSG00000103196", intgroup="dex")
```

```{r}
# Return the data instead of plotting
d <- plotCounts(dds, gene="ENSG00000103196", intgroup="dex", returnData=TRUE)
head(d)
```
use this returned data to plot a boxplot with function boxplot()
```{r}
boxplot(count ~ dex , data=d)
```

We can also make a ggplot for the data 
```{r}
library(ggplot2)
ggplot(d, aes(dex, count)) + geom_boxplot(aes(fill=dex)) + scale_y_log10() + ggtitle("CRISPLD2")
```

# ma & volcano plots 
let’s add a column called sig to our full res results that evaluates to TRUE if padj<0.05, and FALSE if not, and NA if padj is also NA.
```{r}
res$sig <- res$padj<0.05

# How many of each?
table(res$sig)

sum(is.na(res$sig))
```

```{r}
plotMA(res, ylim=c(-2,2))
```
```{r}
resLFC <- lfcShrink(dds, coef=2)
resLFC

plotMA(resLFC, ylim=c(-2,2))
```

```{r}
ggplot(as.data.frame(res), aes(log2FoldChange, -1*log10(pvalue), col=sig)) + 
    geom_point() + 
    ggtitle("Volcano plot")
```


